\(Q\)-Bosons and their relations with integrable hierarchies

### 2024-04-16 draft.tex
remove old files and improve q-bosons

\subsubsection{Partitions representation}
Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by the the conditions 
\(\phi_i\ket{0}_i\). The representation of this Hilbert space, denoted by
\(\mathcal{F}_i\), is given by \(\ket{n_i}_i = \phi_i^\dagger \ket{0}_i\).

Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
\otimes \cdots \otimes  \ket{0}_M\),
the Fock space is defined as 
\begin{equation}
  \mathcal{F} = \bigotimes_{i=0}^M \mathcal{F}_i = 
  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
\end{equation}
where the states \(\ket{\vec{n}}\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (\phi_0^\dagger)^{n_0} (\phi_1^\dagger)^{n_1} \cdots  (\phi_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
We will assume that these states are normalized, that is
\(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).

@@

\subsection{q-Bosons}
The Hamiltonian for the q-boson model is
given by
\begin{equation}
  \mathcal{H} = -\frac{1}{2} \sum_{n=0}^M
  \left(b_n^\dagger b_{n+1} + b_n b_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
\end{equation}
where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and as in
the case of the phase mode, we impose periodic boundary conditions
\(b_{M+1} = b_0\) and \(b_{M+1}^\dagger = b_0^\dagger\). The set of
operators \(\{b_i, b^\dagger,\mathcal{N}_i\}_{i=0}^M\) form the
\(M+1\) copies of independent q-boson algebras:
\begin{equation}
[\mathcal{N}_i, b_j^\dagger]=\delta_{i,j} b_i^\dagger\; , \quad 
[\mathcal{N}_i, b_j]=-\delta_{i,j}b_i\; , \quad
[B_i, B_j^\dagger]= \delta_{i,j} q^{-2\mathcal{N}_i}  \equiv \delta_{i,j} Q^{\mathcal{N}_i}\; , 
\end{equation}
where we write the deformation parameter as \(Q = q^{-2}\). 
In the limit \(Q\to 1\), the q-bosons are ordinary bosons, and 
the limit \(Q\to 0\) (\(q\to \infty\)) gives the phase model we have discussed above.

\subsubsection{Representation}
We build the representation of the q-boson algebra with some modifications of the phase 
model we have just discussed. We suppose that the action of the operators \(\{b_i, b_i^\dagger\}\) 
satisfies 
\begin{subequations}
\begin{alignat}{3}
    & b_0\ket{n_0} =(1 - \delta_{0, n_0}) \frac{1 - Q^{n_0}}{(1 - Q)^{1/2}}\ket{n_0 - 1}
    &\hspace{0.5cm}& b_0^\dagger \ket{n_0} =  \frac{1}{(1 - Q)^{1/2}}\ket{n_0 + 1}  \\
    & b_i\ket{n_i} = \frac{(1 - \delta_{0, n_i})}{(1 - Q)^{1/2}}\ket{n_i - 1}
    && b_i^\dagger \ket{n_0} =  \frac{(1 - Q^{n_i+1})}{(1 - Q)^{1/2}}\ket{n_0 + 1}\quad i\neq 0
\end{alignat}
\end{subequations}
where the modes \(\{b_0, b_0^\dagger \}\) satisfies a slightly different relation.

\subsubsection{Bethe Ansatz}
The \(L\)-operator for the q-boson is given by
\begin{equation}
  L_n(x, Q) =
  \begin{pmatrix}
    x^{-1/2} & (1 - Q)^{1/2} b_n^\dagger \\ (1 - Q)^{1/2} b_n & x^{1/2}
  \end{pmatrix}\; .
\end{equation}
The monodromy matrix is 
\begin{equation}
  T_a(x,Q) = L_{am}(x, q)  \dots  L_{a0}(x, q) = 
  \begin{pmatrix}
    A(x) & B(x) \\ C(x) & D(x)
  \end{pmatrix}\; .
\end{equation}

The eigenstates of the Hamiltonian are of the form
\begin{equation}
  \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N B(y_j, Q) \ket{\bm{0}}\qquad 
  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j, Q) \; .
\end{equation}
When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations 
\begin{equation}
  y^{N + M}_i =\prod_{\substack{j = 1 \\ j \neq i}}^N\frac{Q y_i - y_j}{y_i - Q y_j}\; , \qquad i = 1, \dots, N\; , 
\end{equation}
Once again, we will consider \emph{off-shell} Bethe states.

@@

This section discusses presence of integrable hierarchies in the phase
model.  We argue that it is also a Toda hierachy tau function, and we
discuss some consequences of this fact, in particular, its connection
with a matrix model. Moreover, we also point the presence of
correlation functions in this model that also satisfy the KP hierarchy
equations. We finally argue how these results also imply that these
scalar products are also KP hierarchy tau functions, what is expected from 
the results of Wheeler~\cite{Wheeler:2010vmq}.

@@

From the coordinate expansion of the Bethe
vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
\begin{equation}
\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
 \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
\end{equation}
where \(s_{\mu/\lambda}\) are Schur polynomials~\cite{Macdonald:1998}.
Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
\begin{equation}
\begin{split}
  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
\end{split}
\end{equation}
where we have used that \(\phi^\dagger \ket{\bm{0}} = \ket{(m)}\), and in the second line
we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
This expression also shows that the expresison \(P_m\) can also be written as a tau function.

@@

Observe that these elements have finitely many non zero elements, and
that these include diagonal terms \(\bm{\colon} \psi_i
\psi_i^\star\bm{\colon}\), upper and lower triangular terms. These are
equivalent to the Number operators \(\mathcal{N}\), annihilation
\(\phi\) and creation \(\phi^\dagger\) , respectivelly.  The central
charge is equivalent to the vaccum projection \(\pi =
\ket{\bm{0}}\bra{\bm{0}}\). Therefore, the elements
\begin{equation}
  G \in GL(\infty) \mapsto
  \mathcal{G} = \exp \left(\sum_{i=1}^M \sum_{a=1}^3  c_{i, a} T_i^{a}  + c \pi_i\right)
\end{equation}
\(T^{(a)}_i \in \{ \mathcal{N}_i, \phi_i, \phi_i^\dagger \}_{i=1}^M\).

@@

Putting all these facts together, we have that the tau functions of the Toda 
hierarchies, given by
\begin{equation}
  \tau_s(\bm{x}, \bm{y}) = \bra{s} \prod_{i} \Gamma_+(x_i) G \prod_j \Gamma_+(x_j) \ket{s}\; ,
\end{equation}
are maped into objects of the form 
\begin{equation}
\begin{split}
  \tau_0(\bm{x}, \bm{y}) & = \bra{\Psi(\bm{x})} \mathcal{G} \ket{\Psi(\bm{y})} \\
  \\ & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
  \mathcal{G} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\; ,
\end{split}
\end{equation}
where we necessarily have \(s=0\) in the phase model

@@

\section{Scalar products for the Q-bosons}

Here I would like to start considering the q-boson problem. It has
been shown~\cite{Wheeler:2010vmq, Sulkowski:2008mx, Tsilevich:2006}
that the scalar product of two off-shell Bethe states in the q-boson
model is given by
\begin{equation}
S_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
\prod_{k=1}^N B(y_k, Q) \ket{\bm{0}}
= \sum_{\lambda \subseteq
  (N,N,M)} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
\end{equation}
where \(P_\lambda\) denote the Hall-Littlewood polynomials.

It turns out that this expansion is  also related to the Cauchy identity for 
Hall-Littlewood polynomials~\cite{Macdonald:1998}
\begin{equation}
\sum_{\lambda} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
= \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k}\; ,
\end{equation}
and using the results in~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}, 
we can expand this identity as 
\begin{equation}
\label{eq:cachy-hl}
 \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k} = 
\sum_{\lambda} S_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
\sum_{\lambda} S_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; ,
\end{equation}
where the polynomials, the big-Schur~\cite{Wheeler:2018}, are given by 
a Jacobi-Trudi formula
\begin{equation}
  S_{\lambda} (\bm{y}, Q) = \det(q_{\lambda_i -i + j}(\bm{y}, Q))\; , 
\end{equation}
and the coefficients \(q_m\) can be obtained from the following expression
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_{j} \frac{1-Q y_j z}{1 - y_j z}
\end{equation}
where \(z\) is a formal variable. It has been argued
in~\cite{Foda:2008hn} that if we interpret the \(Q\)-Schur functions
as Plücker coordiantes, the identity~(\ref{eq:cachy-hl}) is a KP-tau
function.

@@

\subsection{Supersymmetric Schur polynomials}

@@

%\input{./sections/fermions.tex}
%
%\input{./sections/schurpol.tex}
%
%\input{./sections/transfer.tex}
%
%\input{./sections/t-fermions.tex}
%
%\input{./sections/top-strings.tex}

%\addcontentsline{toc}{section}{References}

>>>>

\subsubsection{Representation}
Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by \(\phi_i\ket{0}_i
= 0\).  The representation of this Hilbert space, denoted by
\(\mathcal{F}_i\), is given by \(\ket{n_i}_i = \phi_i^\dagger
\ket{0}_i\).

Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
\otimes \cdots \otimes  \ket{0}_M\),
the Fock space is defined as 
\begin{equation}
  \mathcal{F} = \bigotimes_{i=0}^M \mathcal{F}_i = 
  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
\end{equation}
where the states \(\ket{\vec{n}}\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (\phi_0^\dagger)^{n_0} (\phi_1^\dagger)^{n_1} \cdots  (\phi_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
These states are normalized, that is
\(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).

@@

\subsection{q-Bosons}
The Hamiltonian for the q-boson model is
given by
\begin{equation}
  \mathcal{H} = -\frac{1}{2} \sum_{n=0}^M
  \left(b_n^\dagger b_{n+1} + b_n b_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
\end{equation}
where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and as in
the case of the phase mode, we impose periodic boundary conditions
\(b_{M+1} = b_0\) and \(b_{M+1}^\dagger = b_0^\dagger\). The set of
operators \(\{b_i, b^\dagger,\mathcal{N}_i\}_{i=0}^M\) form the
\(M+1\) copies of independent q-boson algebras:
\begin{equation}
[\mathcal{N}_i, b_j^\dagger]=\delta_{i,j} b_i^\dagger\; , \quad 
[\mathcal{N}_i, b_j]=-\delta_{i,j}b_i\; , \quad
[b_i, b_j^\dagger]= \delta_{i,j} q^{-2\mathcal{N}_i}  \equiv \delta_{i,j} Q^{\mathcal{N}_i}\; , 
\end{equation}
where we write the deformation parameter as \(Q = q^{-2}\). 
In the limit \(Q\to 1\), the q-bosons are ordinary bosons, and 
the limit \(Q\to 0\) (\(q\to \infty\)) gives the phase model we have discussed above.

\subsubsection{Representation}
We build the representation of the q-boson algebra with some modifications of the phase 
model we have just discussed. Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by
\(b_i\ket{0}_i = 0\). The representation of this Hilbert space, denoted by
\(\mathcal{F}_i^Q\), is given by \(\ket{n_i}_i \propto b_i^\dagger \ket{0}_i\).

Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
\otimes \cdots \otimes  \ket{0}_M\),
the Fock space is defined as 
\begin{equation}
  \mathcal{F}_{Q} = \bigotimes_{i=0}^M \mathcal{F}_i^Q = 
  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
\end{equation}
where the actions of the operators \(\{b_i, b_i^\dagger\}\) are given by 
the following relations
\begin{subequations}
\begin{alignat}{3}
    & b_0\ket{n_0} =(1 - \delta_{0, n_0}) \frac{(1 - Q^{n_0})}{(1 - Q)^{1/2}}\ket{n_0 - 1}
    &\hspace{0.5cm}& b_0^\dagger \ket{n_0} =  \frac{1}{(1 - Q)^{1/2}}\ket{n_0 + 1}  \\
    & b_i\ket{n_i} = \frac{(1 - \delta_{0, n_i})}{(1 - Q)^{1/2}}\ket{n_i - 1}
    && b_i^\dagger \ket{n_0} =  \frac{(1 - Q^{n_i+1})}{(1 - Q)^{1/2}}\ket{n_0 + 1}\quad i\neq 0\; .
\end{alignat}
\end{subequations}
Observe that the modes \(\{b_0, b_0^\dagger \}\) satisfies a slightly
different relation.

Therefore, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (b_0^\dagger)^{n_0} (b_1^\dagger)^{n_1} \cdots  (b_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
Moreover, the norm of these states satisfy 
\begin{equation}
  \bracket{\vec{n}}{\vec{m}} = \frac{[m_0]!}{\prod_{i=1}^M[m_i]!} \delta_{\vec{n}\vec{m}}\qquad 
    [n]! =
    \left\{
    \begin{array}{ll}
    \prod_{i=1}^n(1 - Q^i) & \textrm{if}\ \ n\neq 0 \\
    1 & \textrm{otherwise}
  \end{array}\right.\; .
\end{equation}
As before, we can associate to a given states \(\ket{\vec{n}} =
\ket{n_0, n_1, \dots, n_M}\), a partition -- or Young diagram --
\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\), but now we assume that 
there is a proportionaly  factor relating these two objects, 
\begin{equation}
  \ket{\lambda}  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
\end{equation}
where \(p_i(\lambda)\) denotes the number of parts of size \(i\) in
the partition. These partition states satisfy \(
\bracket{\lambda}{\mu} = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
again, we see that the correspondence is not unique since the number
of oscilators at the site \(i=\) are completely ignored in the
partition notation.  An finally, if the number of particles is fixed,
then \(n_0 = N - \ell(\lambda)\).

\subsubsection{Bethe Ansatz}
The \(L\)-operator for the q-boson is given by
\begin{equation}
  L_{an}(x, Q) =
  \begin{pmatrix}
    x^{-1/2} & (1 - Q)^{\frac{1}{2}} b_n^\dagger \\ (1 - Q)^{\frac{1}{2}} b_n & x^{1/2}
  \end{pmatrix}_a\; .
\end{equation}
The monodromy matrix is 
\begin{equation}
  T_a(x,Q) = L_{am}(x, Q)  \dots  L_{a0}(x, Q) = 
  \begin{pmatrix}
    A(x, Q) & B(x, Q) \\ C(x, Q) & D(x, Q)
  \end{pmatrix}_a\; .
\end{equation}

The eigenstates of the Hamiltonian are of the form
\begin{equation}
  \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N \mathbb{B}(y_j, Q) \ket{\bm{0}}\qquad 
  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j, Q) \; .
\end{equation}
where \(\mathbb{B}(x, Q) = x^{M/2} B(x, Q)\) and \(\mathbb{C}(x, Q) =
x^{M/2} C(1/x, Q)\).  When the set of parameters \(\{ y_j \ | \ j =1, \dots , N\}\)
satisfies the Bethe equations
\begin{equation}
  y^{N + M}_i =\prod_{\substack{j = 1 \\ j \neq i}}^N\frac{Q y_i - y_j}{y_i - Q y_j}\; , \qquad i = 1, \dots, N\; , 
\end{equation}
we say that the states \(\ket{\Psi(y_1, \dots, y_N; Q)}\) are \emph{on-shell} Bethe states, otherwise, 
we have \emph{off-shell} Bethe states.

@@

This section discusses the presence of integrable hierarchies in the phase
model.  We argue that it is also a Toda hierachy tau function, and we
discuss some consequences of this fact, in particular, its connection
with a matrix model. Moreover, we also point the presence of
correlation functions in this model that also satisfy the KP hierarchy
equations. We finally argue how these results also imply that these
scalar products are also KP hierarchy tau functions, what is expected from 
the results of Wheeler~\cite{Wheeler:2010vmq}.

@@

From the coordinate expansion of the Bethe
vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
\begin{equation}
\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
 \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
\end{equation}
where \(s_{\mu/\lambda}\) are Schur polynomials~\cite{Macdonald:1998}.
Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
\begin{equation}
\begin{split}
  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
\end{split}
\end{equation}
where we have used that \(\phi^\dagger \ket{\bm{0}} = \ket{(m)}\), and in the second line
we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
This expression also shows that the expression \(P_m\) can also be written as a tau function.

@@

Observe that these elements have finitely many non zero elements, and
that these include diagonal terms \(\bm{\colon} \psi_i
\psi_i^\star\bm{\colon}\), upper and lower triangular terms. These are
equivalent to the Number operators \(\mathcal{N}\), annihilation
\(\phi\) and creation \(\phi^\dagger\) , respectivelly.  The central
charge is equivalent to the vaccum projection \(\pi =
\ket{\bm{0}}\bra{\bm{0}}\). Therefore, the elements
\begin{equation}
   GL(\infty) \ni G \mapsto
   \mathcal{G} = \exp \left[\sum_{i=1}^M \left( \sum_{a=1}^3
   c_{i, a} T_i^{a}  + c \pi_i \right)\right]\; ,
\end{equation}
where \(T^{a}_i \in \{ \mathcal{N}_i, \phi_i, \phi_i^\dagger \}_{i=1}^M\).

@@

Putting all these facts together, we have that the tau functions of the Toda 
hierarchies, given by
\begin{equation}
  \tau_s(\bm{x}, \bm{y}) = \bra{s} \prod_{i} \Gamma_+(x_i) G \prod_j \Gamma_+(x_j) \ket{s}\; ,
\end{equation}
are mapped into objects of the form 
\begin{equation}
\begin{split}
  \tau_0(\bm{x}, \bm{y}) & = \bra{\Psi(\bm{x})} \mathcal{G} \ket{\Psi(\bm{y})} \\
  \\ & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
  \mathcal{G} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\; ,
\end{split}
\end{equation}
where we necessarily have \(s=0\) in the phase model

@@

\section{Tau functions in the Q-bosons model}

We now move out attention to the case of the q-bosons. The 
analysis is parallel to what we did before, but the  specific 
details are very different. We start with the study of the norm of  
two off-shell Bethe states.

It has been shown~\cite{Tsilevich:2006} conf. \cite{Sulkowski:2008mx,
  Wheeler:2010vmq}, that the Bethe states \(\ket{\Psi(\bm{x})}\) have 
coordiante expansion given by
\begin{equation}
  \ket{\psi(\bm{x})} = \sum_{\mu \subset [N,M]} P_\mu(\bm{x}, Q) \ket{\mu}  \qquad 
  \bra{\psi(\bm{x})} = \sum_{\mu \subset [N,M]} P_\mu(\bm{x}, Q) \bra{\mu}\; ,
\end{equation}
where \(P_\lambda\) denote the Hall-Littlewood polynomials.  In this
form, scalar product of two off-shell Bethe states in the q-boson
model is easily calculated to be
\begin{equation}
S_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
\prod_{k=1}^N B(y_k, Q) \ket{\bm{0}}
= \sum_{\lambda \subseteq
  (N,N,M)} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
\end{equation}
where we use that \(\bracket{\lambda}{\mu} = b_\lambda(Q)
\delta_{\lambda, \mu}\).  It turns out that this expansion is also
related to the Cauchy identity for Hall-Littlewood
polynomials~\cite{Macdonald:1998}
\begin{equation}
\label{eq:cauchy_hl}
\sum_{\lambda} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
= \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k}\; ,
\end{equation}

Ourt goal is to consider some elementary properties of Hall-Littlewood polynomials 
to understand aspects of this expansion.

\subsection{Scalar product: Big Schur functions}

Let us now consider the results in~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}.
We can expand the Cauchy identity~(\ref{eq:cauchy_hl}) as
\begin{equation}
\label{eq:cachy-hl}
 \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k} = 
\sum_{\lambda} S_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
\sum_{\lambda} S_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; ,
\end{equation}
where the polynomials, the big-Schur~\cite{Wheeler:2018}, are given by 
a Jacobi-Trudi formula
\begin{equation}
  S_{\lambda} (\bm{y}, Q) = \det(q_{\lambda_i -i + j}(\bm{y}, Q))\; , 
\end{equation}
and the coefficients \(q_m\) can be obtained from the following expression
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_{j} \frac{1-Q y_j z}{1 - y_j z}
\end{equation}
where \(z\) is a formal variable. It has been argued
in~\cite{Foda:2008hn} that if we interpret the \(Q\)-Schur functions
as Plücker coordiantes, the identity~(\ref{eq:cachy-hl}) is a KP-tau
function.

@@

\subsection{Scalar product: Supersymmetric Schur polynomials}

@@
